### 2024-10-23 paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage[]{natbib}
\usepackage{textcomp}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%

%% load any required packages here

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{amssymb}
\usepackage{tikz-cd}

\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\hypersetup{
  pdftitle={No More, No Less Than Sum of its Parts: Monoids and The Algebra of Graphics, Statistics, and Interaction},
  pdfkeywords={Interactive data visualization, category theory, abstract
algebra, the Gestalt Principles, descriptive statistics},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf No More, No Less Than Sum of its Parts: Monoids and The
Algebra of Graphics, Statistics, and Interaction}

  \author{
        Adam Bartonicek \\
    Department of Statistics, The University of Auckland\\
     and \\     Simon Urbanek \\
    Department of Statistics, The University of Auckland\\
     and \\     Paul Murrell \\
    Department of Statistics, The University of Auckland\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf No More, No Less Than Sum of its Parts: Monoids and The
Algebra of Graphics, Statistics, and Interaction}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
Interactive data visualization has become a staple of modern data
presentation. Yet, despite its growing popularity, there still remains a
lack of a formal framework for turning raw data into summary statistics
that can then be displayed by interactive graphics. This gap may stem
from a subtle yet profound issue: while we would often like treat
statistics, graphics, and interaction as independent entities, they are
in fact deeply connected. This paper aims to shed light on this
interdependence by relating it to two fundamental concepts from category
theory: groups and monoids. Specifically, if we want our graphics to
support interactive features which split our data into parts and then
combine these parts back together (such as linked selection), then the
statistics in our plots need to posses certain algebraic properties.
Ultimately, by grounding our thinking in these concepts, we may be able
to build more flexible and expressive interactive data visualization
systems.
\end{abstract}

\noindent%
{\it Keywords:} Interactive data visualization, category theory,
abstract algebra, the Gestalt Principles, descriptive statistics

\vfill

\newpage
\spacingset{1.9} % DON'T change the spacing!

\nocite{horst2020} \nocite{magritte1929}

\section{\texorpdfstring{Introduction
\label{sec:intro}}{Introduction }}\label{introduction}

\begin{quote}
``Effective graphical analysis makes things seem obvious, the effort
involved in making the graphical analysis effective is not so obvious.''
\citep{unwin2018}
\end{quote}

The essence of data visualization is comparison. When we visualize, we
split our data into parts, summarize each part by one or more summary
statistics, encode the summaries into visual attributes such as
position, size, or color, and, finally, use these attributes to draw
geometric objects
\citetext{\citealp{bertin1983}; \citealp{wilkinson2012}; \citealp[for a
recent review, see][]{franconeri2021}; \citealp{wilke2019}}. For
example, when drawing a typical barplot, we split our data into parts
based on the levels of some categorical variable, count the number of
cases in each part, and finally represent these counts by drawing bars
of corresponding heights. In this way, each geometric object (bar)
becomes a representation of its own small data set, a subset of the
original data.

To ground this idea in a concrete example that will be used throughout
this paper, imagine going through your bookshelf, sorting the books by
genre, and stacking them on top of each other. You will construct a
physical version of the barplot, forming one ``bar'' (stack) per genre:
one bar for sci-fi novels, one bar for romance, one bar for detective
fiction, and so on. Even this very simple example reveals some
interesting mathematical properties. For example, a single book cannot
go into two stacks (the subsets of the data are \emph{disjoint}) and if
all of the stacks are combined back together, we should end up with the
bookshelf we started with \citep[the subsets are \emph{surjective} and
cover the entirety of the data set, see][]{ziemkiewicz2009}.

Now, just as we can learn about the contents of our library by sorting
books into stacks, we can learn about our data by splitting it into
parts. A natural question then arises: can we learn more by subdividing
our data into even smaller parts? For instance, in our book example, we
can split each stack of books into smaller stacks based on the gender of
the author. Each smaller stack then represents the unique combination
(Cartesian product) of the two categorical variables (genre and gender).
Moreover, there arises a hierarchical relationship between the stacks:
by combining genre-gender stacks (child components), we can recover the
original genre stack (parent component). Similar situation often arises
in interactive graphics, where we use techniques such as linked
selection to highlight specific subsets of our data. Either way, by
forming these finer partitions of our data, we can represent it with
greater granularity and discover trends that may otherwise hide in the
aggregate.

However, once we have subdivided each category into its child
components, we need to represent the information in a way that visually
preserves this hierarchical relationship. In data visualization, there
are two popular methods for doing this, known as ``stacking'' and
``dodging'' (see Figure \ref{fig:playfair}). These methods have existed
at least since the time of William Playfair
\citetext{\citeyear{playfair1801}; \citeyear{playfair1822}}. Let's
briefly illustrate them on the example of the barplot. In a stacked
barplot, bar segments representing the product of the two factor
variables are plotted vertically on top of each other (``stacked''; the
operation can be also thought of as ``highlighting'' parts of the bar).
In contrast, in a dodged barplot, bars are plotted side-by-side, as
``clusters''. Finally, there is also a third more recent method, known
as layering, where we plot bar segments in separate graphical layers and
use partial transparency to mitigate overplotting.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/playfair-1.pdf}
\caption{\label{fig:playfair}Examples of three methods for dealing with
bars that have split into child components (segments). Top row: dodging
(left) and stacking/highlighting (right), both by William Playfair
\citetext{\citeyear{playfair1801}; \citeyear{playfair1822}}. Bottom row:
layering, created with \texttt{ggplot2}
\citetext{\citealp{wickham2010}; \citealp[the source of the underlying
data is the U.S. fuel economy data set,][]{mpg2024}}.}
\end{figure}

Much has been written about the relative merits of stacking, dodging,
and layering. For example, layering is only useful with few categories,
as blending many colors can make it difficult to tell the categories
apart \citep{franconeri2021, wilke2019}. Further, in a landmark study,
\citet{cleveland1984} showed that people tend to be less accurate when
reading information from stacked bar charts as opposed to dodged bar
charts. Specifically, since the lower y-axis coordinate of a stacked
segment is pushed up by the cumulative height of the segments below, it
becomes difficult to accurately compare segments' length, both within
and across bars \citep{cleveland1984}. Subsequent research has
independently validated these findings and expanded upon them \citep[see
e.g.][]{heer2010, thudt2016, quadri2021}. Due to this suboptimal
statistical legibility, many data visualization researchers have urged
caution about stacking \citep[see
e.g.][]{byron2008, cairo2014, franconeri2021}, and some have even
discouraged its use altogether \citep{kosara2016, wilke2019}.

However, despite the mixed reception towards stacking in static
visualization, it remains a popular choice in interactive data
visualization. Among popular interactive data visualization libraries,
stacked plots such as bar charts and histograms make frequent appearance
\citep[see
e.g.][]{cook2023, murray2017, vegalite2022, vanderplas2018, sievert2020, swayne2003, xie2014, theus2002, urbanek2003, urbanek2011, wills2008, ware2019}.
What is the reason for this? Has the interactive data visualization
community simply failed to catch up with the current best practices? Or
is there perhaps more going on?

We argue that not only is stacking a useful technique, despite its
limitations, but that it can also offer valuable insights into the
mathematical structure underlying our visualizations. To make the
strongest case possible, we bring forward evidence from several levels
of the visualization process. We start by discussing the properties of
visual perception and how they apply to both static and interactive
visualizations. Next, we explore the process of computing statistical
summaries that form the basis of our plots. Finally, we dive into the
algebraic properties of the functions underlying these statistical
summaries. Throughout the text, we use the example of the humble barplot
to ground our discussion.

Our goal is to convince you that, in order to produce coherent and
well-behaved interactive visualizations, we simply cannot treat
graphics, statistics, and interaction as independent components. Instead
we need to consider them holistically, within the context of their
respective visual and mathematical properties. Understanding this deep
web of interdependence may in turn empower us to innovate and build new
kinds of robust and expressive interactive visualizations.

\section{Visual Perception and
Interactivity}\label{visual-perception-and-interactivity}

\subsection{Visual perception and the Gestalt
Principles}\label{visual-perception-and-the-gestalt-principles}

Over the course of our evolutionary history, the human brain has evolved
a sophisticated network for processing visual input. This visual
processing pipeline has been the subject of a great body of research in
cognitive neuroscience
\citetext{\citealp{goebel2004}; \citealp{knudsen2020}; \citealp[for a
brief overview, see][20]{ware2019}}, and many of its findings are
directly applicable to data visualization. For example, certain
particularly salient visual stimuli are perceived automatically, rapidly
(\textless10 ms), in parallel, and at a constant speed regardless of the
number of distractors \citep{treisman1985}. This phenomenon, known as
pre-attentive processing, is key for data visualization, as the right
use of visual attributes such as color, contrast, or shape can help us
design figures which communicate insights much more effectively
\citep{ware2019}.

However, many of the visual perception findings are quite specific and
niche; to guide effective visualization, we also need some general,
high-level rules. One such set of high-level rules are the Gestalt
principles of visual perception \citep[see
e.g.][]{cairo2012, ware2019, rosli2015, vanderplas2020, todorovic2008, pinker1990}.
These principles formulate a set of fundamental ``laws'' related to the
perception of (primarily visual) patterns. They were discovered by the
founders of the Gestalt school of psychology early in the 20th century,
primarily by means of self-observation, and, despite not being grounded
in sound understanding of the neural mechanisms of the brain, have
nevertheless been later independently supported by findings from
cognitive neuroscience \citep{guberman2017, ware2019, wertheimer1938}.

Among the Gestalt principles, there are a few which are particularly
relevant to data visualization. First, there are the two principles of
proximity and similarity. These principles state that objects that which
appear in close together and/or similar in appearance tend to be
perceived as part of single group. Second, the closure and common region
principles state that regions with closed contours tend to be perceived
as single, unified objects. Finally, the figure and ground principle
states that we tend to perceive one part of a visual scene as the
central object (the ``figure'') and the rest as background (the
``ground''). This can be modulated by other Gestalt principle: for
example, smaller, darker, and symmetric objects with closed contours are
more likely to be perceived as the figure \citep{ware2019}.

Whether we intend them to or not, the Gestalt principles affect how our
data visualizations are perceived. As such, it is a good idea to try to
use them to our advantage. Sometimes, little effort is needed. Such is
the case, for example, when identifying clusters in a scatterplot
\citep{rosli2015}: the principles of proximity and similarity ensure
that clouds of similarly shaped and colored points will be perceived as
discrete groups. In other cases, however, our design choices can have a
significant impact. For instance, another way to make objects be
perceived as part of a group is to draw a closed shape around them
\citep{cairo2012}. Alternatively, if we want to highlight a particular
part of a data set, we can leverage the figure and ground principle and
use attributes such as color and saturation to make it stand out
\citep[for specific examples, see][34]{franconeri2021, wilke2019}.

Finally, ignoring or violating the Gestalt principles can make our
graphics less legible or even misleading. \citet{tufte2001} had
documented many such violations, such as deceptive 3D effects,
inadvertent optical illusions caused by dense stripe patterns, and the
overuse of superfluous graphical elements termed ``chartjunk''.
Similarly, \citet{cairo2019} highlights many cases of graphics that
violate the Gestalt principles. Finally, there is also the issue of
well-designed graphics that communicate no statistical information at
all, or even information that is blatantly false
\citep{cairo2014, franconeri2021, cairo2019, gelman2013, tufte2001}.
Whether intentional or not, poor design can have a significant impact on
the legibility and effectiveness of statistical graphics.

\subsection{Gestalt and Interactivity}\label{gestalt-and-interactivity}

The Gestalt principles apply to both static and interactive
visualizations, however, the way they apply is different. Static data
visualizations can be designed well or poorly, but this design quality
does not change over time. In contrast, interactive visualizations
change over time, with user interaction. As such, an interactive
visualization may conform to the Gestalt principles at one instance of
time and violate them in another. In fact, interactive visualizations
can even differ in how well they leverage design principles
\emph{across} time: the animations that interactions induce may
themselves conform to good or poor design
\citep[see][]{hullman2013, ware2019}.

Thus, interactivity inherently places additional constraints the design
of our figures. If we want well-designed interactive figures, we need to
ensure that they conform to good design principles across all states
induced by user-interaction. This may not always be easy. We will
investigate the design challenged posed by interaction in the following
section, on the example of linked selection.

\subsection{Case study: Linked
Selection}\label{case-study-linked-selection}

Linked selection, also known as linked brushing or linked highlighting,
has been consistently ranked as one of the most useful interactive
features in data visualization \citep[see
e.g.][]{buja1996, heer2012, wilhelm2003, ware2019, ward2015}. By
clicking or clicking-and-dragging over objects in one plot, users can
highlight parts of objects corresponding to the selected cases across
all other ``linked'' plots. This allows the user to rapidly
``drill-down'' \citep{dix1998, theus2002} and explore trends across
different dynamically-generated subsets of the data \citep[similar to
interactive filtering, see e.g.][]{heer2012}. The ability to quickly
materialize different views of the data makes linked selection a
versatile tool for data exploration.

Yet, despite its usefulness, linked selection also imposes some
fundamental constraints on the figure. Specifically, all objects we
visualize need to be able to:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Be selected}
\item
  \textbf{Represent selection}
\item
  (both in a visually and statistically sound way).
\end{enumerate}

While these constraints may not seem too restrictive, we argue that they
pose significant challenges for certain visualization styles.
Specifically, some kinds of plots may produce undesirable visual
behavior when combined with linked selection, and this may in fact
explain why stacking is preferred over dodging within the interactive
data visualization community. Let's explore this idea in more depth by
comparing linked selection with a stacked vs.~dodged barplot.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/display-1.pdf}
\caption{\label{fig:display}The visual coherence of stacking vs.~dodging
during linked selection. Dark segments represent selected cases, light
segments represent unselected cases, and the plots' contour is shown via
a light outline. Plots left to right show more cases being selected. Top
row: the contour of a stacked barplot remains constant throughout
selection (three bars of the same height). Bottom row: the contour of a
dodged barplot changes dramatically with selection. Note the changes to
the upper y-axis limit and the absence of the lighter segment among the
bar representing category B in the rightmost plot (the darker segment
has been stretched to fill the available space; \texttt{ggplot2} default
for handling empty subcategories). The source of the underlying data is
the \texttt{mtcars} dataset \citep{henderson1981}.}
\end{figure}

First, we propose that stacking is the superior method for
\emph{displaying} selection (see Figure \ref{fig:display}). Thanks to
the four Gestalt principles of proximity, similarity, closure, and
common region, in a stacked barplot, each bar presents itself as a
single visual entity. In other words, since the stacked segments are
placed right on top of each other, and share width and a closed border,
they are perceived as part of a unified whole (the bar). This remains
true throughout selection - no matter how the heights of the highlighted
segments change, the height of the whole stacked bar remains constant,
and so does the overall outline of the plot.

Dodged barplots lack this degree of visual consistency. In a dodged
barplot, bar segments placed side-by-side in clusters, and, since each
segment can be of different length, these clusters lack a shared border
(common region). This leads to several design challenges. First, when
selection occurs, the jagged contour of the dodged barplot can change
dramatically as cases are added or removed. This creates distracting
visual clutter that makes it difficult to track changes over time.
Second, since the heights of the bar segments are not bound by any value
present in the plot (unlike the heights of the segments in the stacked
barplot, which are bound by the height of the tallest bar), it is not
clear how to handle the upper y-axis limit. On one hand, we could make
the upper y-axis limit dynamic and have it change in response to
selection. If we do this, we ensure that the bar segments always fit the
plotting region, however, we also lose the context that the limit
provides each time selection occurs. Alternatively, we can set the limit
to a static value, however, then we are at the risk of the bar segments
growing outside of the bounds of the plotting region.

Stacked bars may offer several other advantages over dodged bars within
the context of linked selection. First, we propose that stacking also
provides a simpler mental model for \emph{selection}. Since each stacked
bar is a single visual entity, it makes intuitive sense that clicking it
will select the entire corresponding category. Indeed, this is how many
interactive data visualization systems implement default linked
selection: clicking on part of an object selects all cases corresponding
to the whole object \citep[see
e.g.][]{cook2023, vegalite2022, vanderplas2018, xie2014, theus2002, urbanek2003, urbanek2011, swayne2003}.
On the other hand, with dodged bars, since the segments do not form a
unified whole, the expectation is less clear: when we click on a segment
in a dodged barplot, do we select the entire category or just the cases
corresponding to the individual segment? Second, stacked barplots
maintain constant bar widths and inter-bar gaps throughout selection. In
a dodged barplot, it is not clear what to do when some selection groups
are empty. Should we leave gaps for the (absent) highlighted segments,
leading to asymmetric gaps and potential illusory clustering? Or should
we draw wide segments and only shrink them in response to selection,
avoiding asymmetry but violating the principle of area proportionality
\citep{cleveland1985}? Finally, by presenting fewer visual entities on
the screen, stacked bars may also reduce cognitive load
\citetext{\citealp{sweller2019}; \citealp[for a non-technical overview,
see][]{knaflic2015}}.

To summarize, stacked bars behave in a fairly graceful and visually
pleasing way when combined with selection. Dodged bars, on the other
hand, lack a bounded uniform outline, and this makes them behave
erratically when combined with selection, hurting the interpretability
of the plot. These difference may explain the continued popularity of
stacking within the interactive data visualization community.

However, a key question remains: when is stacking appropriate? So far,
we have focused on the visual aspect of stacking. But stacking goes
beyond aesthetics. This will be the subject of the next section.

\section{Statistics}\label{statistics}

\subsection{From Data to
Representation}\label{from-data-to-representation}

Data visualizations are made up of geometric objects such as points,
lines, or areas. However, these objects alone hold no meaning. Without
context, an image filled with coloured rectangles is just that - an
image (see Figure \ref{fig:magritte}). To turn the into a plot, we need
to make sure that each rectangle accurately represents some underlying
data.

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/magritte-1}

}

\caption{The representation of a thing is not the thing itself. Left: The Treachery of Images by René Magritte [-@magritte1929] makes the point that a painting of a pipe is not a pipe. Right: similarly, an image filled with coloured rectangles is not a barplot, unless the rectangles accurately represent some underlying data.}\label{fig:magritte}
\end{figure}

Thus, we always have to start with the data. But the raw data is rarely
something we can just ``throw'' at the computer screen. Instead, we
often first need to massage and translate it into a format that can be
understood by the graphics device. In most data visualization packages,
this is done in two steps: aggregation and scaling.

First, the data needs to be aggregated or ``wrangled'' into sets of
statistical summaries. These summaries may be very simple: for instance,
in a typical scatterplot, we assign each point the raw values of the x-
and y-variables (the summary function is then just identity,
\(f(x) = x\)). More often, however, the data is grouped and aggregated
in some way. For example, as was discussed previously, in a barplot,
data points are grouped by the levels of the x-axis variable, and each
group is then summarized by some descriptive statistic, such as count,
sum, mean, or maximum. Likewise, in a typical histogram, the x-axis
variable is divided into bins, and each bin is then summarized by the
number of cases that fall inside it.

Second, the computed statistics have to be encoded into graphical
attributes such as position, length, size, or colour \citep[see
e.g.][]{franconeri2021, wilkinson2012, pinker1990}. This is the job of
specialized functions known as scales or coordinate systems \citep[see
e.g.][]{murrell2005, murrell2007, wickham2010, wickham2016, wilkinson2012}.
For instance, in a typical scatterplot, the values of the x-axis
variable are mapped to the x-position, such that points corresponding to
small values of the x-axis variable are placed near the left-hand side
of the plot (and larger values are place near the right-hand side).
These mappings may be subject to non-linear transformations, such as log
or square-root, and may also translate discrete values, such as barplot
categories. Scales and coordinate systems are a rich subject and we have
barely scratched the surface here; interested reader is advised to see
\citet{wilkinson2012}, \citet{ziemkiewicz2009}, or \citet{hotz2020}.

\subsection{Combining Statistics}\label{combining-statistics}

\begin{quote}
``This system cannot produce a meaningless graphic, however. This is a
strong claim, vulnerable to a single counter-example. It is a claim
based on the formal rules of the system, however, not on the evaluation
of specific graphics it may produce.''

``Some of the combinations of graphs and statistical methods may be
degenerate or bizarre, but there is no moral reason to restrict them.''

\citet{wilkinson2012}, The Grammar of Graphics, pp.~15 and 112.
\end{quote}

But what does stacking mean in the context of the statistics we are
trying to represent? To a data visualization novice, stacking may appear
as a purely graphical operation. If they attempt to implement a data
visualization pipeline from scratch, they may leave stacking to the very
end, after the statistical summaries have already been translated into
screen coordinates.

The view of stacking as a purely graphical operation has its appeal.
Many modern data visualization systems follow the grammar-based model of
visualization
\citep{mcnutt2022, kim2022, vanderplas2020, wickham2010, satyanarayan2014, satyanarayan2016, wilkinson2012}.
This model emphasizes compositionality, and enables the user to build
plots from independent, modular components. For example, in the popular
\texttt{ggplot2} library \citep{wickham2010}, plots are built by
combining components such as geometric objects (called \texttt{geoms})
and statistical summaries (\texttt{stats}). This grammar-based approach
offers a great deal of flexibility and expressiveness, and has gained
widespread popularity as a result.

Thus, it may be tempting to treat stacking just as another modular
component of a visualization, such that we may be able to freely stack
any geometric objects and statistics we like. Unfortunately, there is a
deep flaw in this approach. Many data visualization experts have warned
about this:

\begin{quote}
``Stacking is useful when the sum of the amounts represented by the
individual stacked bars is in itself a meaningful amount''
\citep[52]{wilke2019}.
\end{quote}

\begin{quote}
``Because this gives the visual impression of one element that is the
sum of several others, it is very important that if the element's size
is used to display a statistic, then that statistic must be summable.
Stacking bars that represent counts, sums, or percentages are fine, but
a stacked bar chart where bars show average values is generally
meaningless.'' \citep[112]{wills2011}.
\end{quote}

\begin{quote}
``{[}\ldots{]} We do this to ensure that aggregate statistics are always
computed over the input data, and so users do not inadvertantly compute
e.g., averages of averages, which can easily lead to
misinterpretation.'' \citep{wu2022}
\end{quote}

The message is clear: stacking some statistics such as counts, sums, and
percentages is acceptable, but stacking others is not. As outlined by
Wills, while the sum of grouped counts produces a valid overall count,
the sum of group means does not yield a meaningful statistic. Even
averaging group means, as noted by \citet{wu2022}, does not produce a
valid summary statistic - the mean of group means differs from the grand
mean (unless by coincidence). This problem is general and applies beyond
barplots; for example, concerns about the statistical validity of
stacked density plots have also been recently raised \citep{pu2020}.

What exactly makes stacking counts and sums acceptable but stacking
other statistics such as means not? Wilke, Wills, and Wu et al.~all
argue that the aggregated value should be ``meaningful''. In the case of
sums, there is perhaps a colloquial way to put this: ``the whole is
equal to the sum of its parts''. Or, more specifically:

\begin{quote}
``the \emph{sum} of the \emph{sums} on the parts is equal to the
\emph{sum} of all values''
\end{quote}

But what if we replace word ``sum'' with some placeholder (such as
``foo''):

\begin{quote}
``the \emph{foo} of \emph{foos} on the parts is equal to the \emph{foo}
of all values''
\end{quote}

Then, we find there are other mathematical operators that satisfy this
condition. For example, the product of products or the maximum of
maximums are also valid overall statistics, see Figure
\ref{fig:sumprodmax}. In fact, there is an entire class of mathematical
objects with this exact behavior: monoids. We will explore monoids in
the following section.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/sumprodmax-1.pdf}
\caption{\label{fig:sumprodmax} Different operators can be used for
stacking. The values \texttt{1.2}, \texttt{1.8}, \texttt{NA}, and
\texttt{1.5} were assigned to a 2-by-2 table and stacked using three
different operators. Left: stacking using the sum operator. Middle:
stacking using the product operator (note that the y-axis starts at 1).
Right: stacking using the max operator.}
\end{figure}

\subsection{\texorpdfstring{Monoids
\label{sec:monoids}}{Monoids }}\label{monoids}

Monoids are a simple yet powerful concept that plays a key role in
several areas of mathematics, including abstract algebra and category
theory. The short treatment here draws primarily from \citet{fong2019},
\citet{lawvere2009}, \citet{baez2023}, and \citet{milewski2018}. For a
particularly accessible introduction, readers are encouraged to seek out
\citet{fong2019} or \citet{lawvere2009}.

While the term ``monoid'' may sound intimidating, the concept is really
quite simple. A monoid represents a ``whole equal to the sum of its
parts'', if we relax our notion of what it means to ``sum''. More
formally, a monoid is a tuple \((M, e, \otimes)\) consisting of:

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\tightlist
\item
  A collection (set) of objects \(M\)
\item
  A neutral element \(e \in M\) called the \emph{monoidal unit}
\item
  A binary operation \(\otimes: M \times M \to M\) called the
  \emph{monoidal product}
\end{enumerate}

Conforming to two rules:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \emph{Unitality}: \(x \otimes e = e \otimes x = x\)
\item
  \emph{Associativity}:
  \(x \otimes (y \otimes z) = (x \otimes y) \otimes z = x \otimes y \otimes z\)
\end{enumerate}

(for all \(x, y, z \in M\))

In plain words, when we have a monoid, we have some set of elements
(\(m \in M\)), and a way to combine these elements (\(\otimes\)), such
that the order in which we do the combining does not matter
(associativity). Further, among the elements, we also have a special
element (\(e\)), called the ``neutral'' or ``identity'' element, that,
when combined with any other element, does nothing. This identity
element is always unique \citep[the proof is straightforward, for
reference see e.g.][]{fong2019, lawvere2009}.

Let's return to the book example from Section \ref{sec:intro}. Does it
represent a monoid? First of all, what is our set \(M\)? This could be
the set of all book stacks we can form by combining books in our
bookshelf, including an ``empty'' stack of zero books, ``short'' stacks
made up of a single book each, and the tallest stack formed by taking
all of the books in the bookshelf. Second, do we have a neutral element?
Sure, this is the empty stack of zero books. Third, what is our binary
operation? This is just the operation of taking one stack of books and
putting it on top of another.

Now we just need to verify that our example conforms to the two rules.
First, is the operation associative? Yes, since a stack of detective
novels will be the same height regardless of whether we put the novels
by Agatha Christie on top of Jo Nesbø books or vice versa. Second, is
the operation unital? Also yes, since adding zero books on top of a
stack does not change its height. Therefore, we do have a monoid.

We can describe the book example a bit more formally. Specifically,
summation on natural numbers \((\mathbb{N}, 0, +)\) is a typical example
of a monoid, because, as we have shown, it is associative and unital.
Other typical examples of monoids include the previously mentioned
products of real numbers \((\mathbb{R}, 1, \times)\), and the min and
max operators \((\mathbb{R}, \infty, \min)\) and
\((\mathbb{R}, -\infty, \max)\). As a counterexample, exponentiation is
not a monoid, since it is not associative: \(x^{(y^z)} \neq (x^y)^z\).

The definition of a monoid is quite broad however, and encompasses more
exotic structures than just operations on numbers. For example, the set
of booleans \(\mathbb{B} = \{ True, False \}\) equipped with either of
the logical operators \(\mathbf{AND}\) or \(\mathbf{OR}\) is also a
monoid, and so is the multiplication of \(n \times n\) square matrices.
Likewise, the operation of concatenating vectors and computing the
euclidean distance is also a monoid \citep{stepanov2009}:

\[|| \big( || (x, y) ||_2, z \big) ||_2 = \sqrt{ \bigg(\sqrt{x^2 + y^2} \bigg)^2 + z^2} = \sqrt{x^2 + y^2 + z^2} = ||(x, y, z)||_2\]

Even very ``non-number-like'' things such as strings can be a monoid:

\[\texttt{"hello"} + \texttt{""} = \texttt{""} + \texttt{"hello"} = \texttt{"hello"}\]
\[(\texttt{"quick"} + \texttt{"brown"}) + \texttt{"fox"} = \texttt{"quick"} + (\texttt{"brown"} + \texttt{"fox"}) = \texttt{"quick brown fox"}\]

\subsection{Groups}\label{groups}

Groups are another fundamental concept in mathematics, representing the
idea of symmetry and reversible transformations. While introducing two
algebraic concepts in quick succession may seem like a big ask, a group
is essentially just a monoid with one additional property. Specifically,
for all \(x, y, z \in M\), we have:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  \emph{Inverse operator}: There exists \(\otimes^{-1}\) such that if
  \(x \otimes y = z\) then \(z \otimes^{-1} x = y\)
\end{enumerate}

In simple terms, a group embodies the idea that ``the whole is equal to
the sum of its parts, \emph{in a non-destructive way}''. That is, when
we combine two elements, we can always recover either one by
``subtracting'' the contribution of the other. Of note, like the
identity element for monoids, the inverse operator in a group is also
always unique \citep[see e.g.][]{fong2019, lawvere2009}.

Turning back to the book example, we can remove Agatha Christie novels
off the detective novel stack and recover the height of the rest of the
books. Thus, summation on natural numbers is a group:
\(2 + 3 = 5 \implies 5 - 2 = 3\) (technically, we now have to identify
the underlying set as the set of integers \(\mathbb{Z}\)). But, for
example, the min and max operators lack an inverse and so do not
constitute groups: if \(\max{(x, 6)} = 6\), then there is no way to
recover \(x\) from the combined result (\(6\)) alone.

\subsection{Groups, Monoids, and
Stacking}\label{groups-monoids-and-stacking}

We now arrive at the core idea of the present paper. Suppose that we
have some data set \(D\), a way of summarizing this data set via an
associative binary operation \(\otimes\), and a neutral value with
respect to this operation \(e\). In other words, \((D, \otimes, e)\)
forms a monoid. Define \(F(A)\) as the result of summarizing or
``folding'' some subset \(A \subseteq D\) using the binary operation:

\[
F(A) = a_1 \otimes a_2 \otimes a_3 \ldots \otimes a_n
\]

Note that we could have written this as
\(F(A) = (a_1 \otimes (a_2 \otimes (a_3 \otimes ( \ldots \otimes a_n ))))\),
however, since the operation is associative, we do not have to worry
about brackets. Also, if \(A\) is empty, then clearly:

\[F(\varnothing) = e\]

Now comes the key concept: if the conditions above hold, then combining
summaries of two disjoint subsets of the data set gives us the same
result as summarizing the union of the original subsets:

\[
\begin{aligned}
F(A) \otimes F(B) &= (a_1 \otimes a_2 \otimes \ldots a_n) \otimes (b_1 \otimes b_2 \otimes \ldots b_n) \\
&= a_1 \otimes a_2 \otimes \ldots \otimes a_n \otimes b_1 \otimes b_2 \otimes \ldots \otimes b_n \qquad \text{(by associativity)} \\
&= F(A \cup B)
\end{aligned}
\]

Another way to show this is the following commutative diagram:

\[
% https://q.uiver.app/#q=WzAsOCxbMSwwLCJBIFxcY3VwIEIiXSxbMCwxLCJBIl0sWzIsMSwiQSBcXGN1cCBCIFxcY3VwIEMiXSxbMSwzLCJGKEEgXFxjdXAgQikiXSxbMCw0LCJGKEEpIl0sWzIsNCwiRihBIFxcY3VwIEIgXFxjdXAgQykiXSxbMiw2XSxbMSw0XSxbMCwyLCItXFxjdXAgQyJdLFswLDMsIiIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImRvdHRlZCJ9fX1dLFsxLDIsInstIFxcY3VwIChCIFxcY3VwIEMpfSIsMl0sWzEsNCwiIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiZG90dGVkIn19fV0sWzIsNSwiIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiZG90dGVkIn19fV0sWzMsNSwiey0gXFxvdGltZXMgRihDfSJdLFs0LDUsInstIFxcb3RpbWVzIEYoQSBcXGN1cCBDKX0iLDJdLFsxLDAsIi0gXFxjdXAgQiJdLFs0LDMsIi0gXFxvdGltZXMgRihCKSJdXQ==
\begin{tikzcd}
    & A \\
    {A \cup B} && {A \cup B \cup C} \\
    & {F(A)} \\
    {F(A \cup B)} && {F(A \cup B \cup C)}
    \arrow["{- \cup B}"', from=1-2, to=2-1]
    \arrow["{- \cup (B \cup C)}", from=1-2, to=2-3]
    \arrow[dashed, from=1-2, to=3-2]
    \arrow["{- \cup C}"', near start, from=2-1, to=2-3]
    \arrow[dashed, from=2-1, to=4-1]
    \arrow[dashed, from=2-3, to=4-3]
    \arrow["{- \otimes F(B)}"', near start, from=3-2, to=4-1]
    \arrow["{- \otimes F(B \cup C)}", near start, from=3-2, to=4-3]
    \arrow["{- \otimes F(C)}"', from=4-1, to=4-3]
\end{tikzcd}
\] The top and bottom triangles in the diagram both show associativity:
in the top triangle, we can either first take the union of \(A\) and
\(B\) and then union the result with \(C\), or immediately union \(A\)
with \(B \cup C\), and the result will be identical. The same applies
for the monoidal summary in the bottom triangle. By saying that the
diagram commutes, we mean that any two parallel paths (the same start
and endpoint) always produce the same result. Practically, this means we
can either first summarize each set and then combine the summaries,
first union the sets and then summarize the union, or do some mix of
both, and we always end up with a valid summary. As a sidenote, the
diagram above actually corresponds to something called a functor or
monoid homomorphism, however, that is beyond the scope of the present
paper \citep[interested reader should see][]{fong2019, lawvere2009}.

But what we have now is precisely the idea of a stackable statistic. If
our statistic conforms to the diagram, then the combined summary of
subsets of the data is always a valid summary of the whole data set. In
Wilke's words, we have discovered what makes the combined summary a
``meaningful amount''!

It might seem that we are done. However, there is still one subtle issue
left. Suppose we summarize three parts of the data \(A\), \(B\), and
\(C\), and combine these summaries by repeatedly applying the operation,
as a kind of running total \(F^*\):

\[
\begin{aligned}
F^* \big( F(A), F(B), F(C) \big) &= \big( F(A), F(A) \otimes F(B), F(A) \otimes F(B) \otimes F(C) \big) \\
&= \big( F(A), F(A \cup B), F(A \cup B \cup C) \big)
\end{aligned}
\]

Notice that we end up with summaries on nested subsets of the data,
i.e.~\(F(A)\), \(F(A \cup B)\), and \(F(A \cup B \cup C)\). This is
fine, for example, when we do linked selection with a single highlighted
group and want to compare the summary on the highlighted subset (\(A\))
vs.~the entirety the data (\(A \cup B\)). The resulting statistics give
us what we need.

However, what if we want to compare the subsets \(A\) and \(B\)
directly? For example, suppose we do linked selection and with two
highlighted groups (\(A\) and \(B\)). Unfortunately, the properties of
monoids do not guarantee that we will be able to recover \(F(B)\) from
\(F(A \cup B)\) using \(F(A)\). Put plainly, the monoidal operation may
``collapse'' the information contained in the disjoint parts, see Figure
\ref{fig:summax}. To ensure that \(F(B)\) is recoverable from
\(F(A \cup B)\), we also need the inverse operator \(\otimes^{-1}\):

\[F(B) = F(A \cup B) \otimes^{-1} F(A)\]

Importantly, even when the operation does not have an inverse,
\(F(A) \otimes F(B)\) is still a valid summary of \(A \cup B\). It is
just that \(F(A) \otimes F(B)\) may ``forget'' the information contained
in either of its two constituent parts. Note also that this is not a
problem of different permutations of parts producing different sequences
of accumulated summaries - this will happen regardless of whether there
is an inverse or not.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/summax-1.pdf}
\caption{\label{fig:summax} Monoids can ``collapse'' information about
disjoint subsets. Top: since sum is a group with an inverse (minus), we
can recover the value of each part by subtracting the heights of the
preceding segments (i.e.~the part order does not matter). Bottom: since
maximum lacks an inverse, we may lose information about the individual
parts under certain permutations (note the absence of segments in the
middle and rightmost plot).}
\end{figure}

To summarize, whether our statistic is a monoid or a group determines
what kinds of comparisons we can make with stacked objects:

\begin{itemize}
\tightlist
\item
  Monoids: comparing nested subsets, part vs.~whole comparisons
\item
  Groups: comparing disjoint subsets, part vs.~part comparisons
\end{itemize}

This becomes important when implementing interactive data visualization
features such as linked selection. Specifically, if we want to represent
a single highlighted group vs.~the rest of the data via stacked objects,
then it's enough for our summary statistic to be a monoid. If we want to
represent multiple highlighted groups, then our statistic needs to have
an inverse as well. If we fail either of these conditions, we may
produce representations which are either inaccurate (in the case of
non-monoidal statistics), or accurate but not ``full'' (in the case of
monoids but no inverse).

\subsection{Groups, Monoids, and
Graphics}\label{groups-monoids-and-graphics}

We can use the connection between summary statistics and algebraic
structures to offload some of the cognitive effort required when
thinking about new types of interactive visualizations. For example,
while not generally thought of as a summary statistic, the convex hull
\citep[see e.g.][]{barber1996} provides a way to reduce a set of data
points into a simpler ``summary'', see Figure \ref{fig:convexhull}. We
can recast convex hull algorithm as a monoid: starting with an empty
set, we iterate through the data, adding points if they lie outside of
the current hull (and removing any points in the interior of the new
hull). This may not be the most computationally efficient method of
finding convex hulls, however, since it does meet the definition of a
monoid, this tells us \emph{what convex hulls can be used for}. If we
combine two convex hulls, we know that the new hull will be a valid
summary for the union of the underlying data points. In turn, this
allows us to implement interactive features such as single-group
highlighting/selection, safe in the knowledge that the underlying
structure is preserved.

However, does the operation of adding a point to a convex hull have an
inverse? No.~Once we add a point to the hull, we cannot recover the old
hull by ``subtracting'' that point - we do not know how many points from
the interior of the old hull we removed. This tells us that merging two
convex hulls may irrevocably collapse information. As such, we cannot
safely use convex hulls to display multiple selection groups, without
resorting to some ad-hoc methods.

While it is possible to reach the same conclusions via ad hoc reasoning
or manual testing, these methods may be time-consuming and have to be
applied to each new visualization style separately. By grounding our
understanding in the simple theory of groups and monoids, we can bypass
most of this effort and reason about the properties of the visualization
directly.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:convexhull}Convex hull can be used as a valid
representation of single-group selection but not multi-group selection.
The light polygons shows the convex hull of the whole dataset, the dark
polygon shows the hull of one selection group, the white dashed line
shows the hull of a second selection group, and the medium-brightness
polygon shows the combined hull of the two selection groups. Points show
individual cases. Left: since convex hull is a monoid, we can use it to
show a single highlighted group. Right: because convex hull lacks an
inverse, after combining two hulls, we cannot guarantee we will be able
to recover either of the original hulls - we do not know how many
interior points were ``forgotten''. The source of the underlying data is
the Palmer penguins data set \citep{horst2020}.}
\end{figure}

\subsection{\texorpdfstring{Groups, Monoids, and the Neutral Element
\label{sec:advantages}}{Groups, Monoids, and the Neutral Element }}\label{groups-monoids-and-the-neutral-element}

There are other benefits that groups and monoids offer for visualizing
data. One of these is the existence of the neutral element \(e\) (see
Section \ref{sec:monoids}). While seemingly inconspicuous, the neutral
element can actually be quite useful.

Specifically, the neutral element ensures that we can always represent
empty subsets of our data. Turning once more to the book example from
Section \ref{sec:intro}, suppose we split the books into stacks based on
the book's genre and the author's gender, and then summarize each stack
by its total and average word-count. What if there are no books by male
sci-fi authors? For total word-count, we have a meaningful default value
- zero - as in, there are zero words across the zero books by male
sci-fi authors. However, what about the average word-count? Here we run
into trouble, since the mean of an empty set is not defined. It simply
does not make sense to talk about the ``average word-count across zero
books''.

A common approach is to omit the objects representing empty subsets.
However, with non-monoidal summaries, this approach is flawed and leads
to ambiguity. Suppose we are drawing the bars of averages. If we omit
bars representing empty categories, then the missingness of a bar can
indicate one of two things: either there are \emph{no} cases in the
category, or there \emph{are} cases and their average is equal to the
lower y-axis limit. There is just no way to tell from the graphic alone.
In contrast, with monoidal summaries such as sums, the absence of an
object (bar) sends a clear message. Regardless of whether there are zero
or hundred cases in a category, if the corresponding bar is missing,
then we know their sum is zero.

The neutral element can also provide a new perspective on the
longstanding debate in data visualization: should the base of a barplot
always start at zero? Some authors argue that starting from a non-zero
base renders bar lengths meaningless \citep{cleveland1985}. On the other
hand, \citet{wilkinson2012}, for example, suggests that the bar base is
really a statement about scales rather than graphics, and as such it
does not have to be zero. We offer the following new perspective: often,
it might make sense to start the base of the barplot at \(e\), the
neutral element, since this represents the default state of ``doing
nothing''. In other words, the bar base is tied to the properties of the
\emph{statistic}, rather than just the graphic or the scale. For
instance, in a logarithmic barplot, the underlying monoidal operation is
multiplication, and so the base of the bars should start at one.

These example provides further support for the claim that we cannot
separate statistics from other components such as graphics and scales.
Even before we apply any transformations, the method we have used to
derive our summary statistics determines important properties of our
visualization. We need to be mindful of this.

\subsection{Groups, Monoids, and
Computation}\label{groups-monoids-and-computation}

Groups and monoids also offer some inherent computational advantages.
Specifically, if a summary statistic is a monoid, associativity
guarantees that it can be computed in a single pass through the data
(i.e.~\(O(n)\) complexity), and in a distributed/parallel fashion
\citep[see e.g.][]{lammel2008, lin2013}. This is makes monoids popular
in functional programming, where they are known as the \texttt{fold} or
\texttt{reduce} functions \citep[see e.g.][]{milewski2018, abelson2022}.

While computational efficiency is usually not the number one concern in
static visualization, it becomes significant in the context of
interactive graphics. Here, we may be able to use the properties of
groups and monoids to avoid unnecessary computation. Specifically, when
using groups and monoids, we can exploit associativity: the fact that
summaries on fine-grained subsets of the data ``add up'' to the
summaries on coarse-grained subsets.

Consider a stacked histogram, with linked selection, reactive axes, and
interactive manipulation of binwidth and anchor. When the user engages
in selection, the counts within the individual highlighted segments may
change but the counts within the whole bars will stay constant.
Therefore, there is no need to update the upper y-axis limit. However,
if the user updates either the binwidth or the anchor, then both the
counts within the whole bars \emph{and} within the highlighted segments
may change, and so we need to update the graphic appropriately.

In other words, associativity allows us to implement strategic updating.
Fine-grained summaries (individual segments) need to be updated whenever
coarse-grained summaries (whole stacked bars) change, but not vice
versa. By pre-computing summaries on different levels of aggregation, we
may be able to reduce computational effort.

In fact, if our summary statistic has an inverse, we can be even more
computationally efficient. Suppose that one data point out of a thousand
is removed from selection. If our summary has an inverse, we can use it
to subtract the contribution of that single data point, instead of
having to recompute the whole summary from scratch. This is, in fact,
the principle behind the fast updates in the popular JavaScript library
Crossfilter, which is unfortunately no longer actively maintained
\citetext{\citealp{crossfilter2023}; \citealp[see also the Github
discussion initiated by][]{crossfilter2024}}\footnote{Note that, as the
  present date, we are not aware of any explicit discussion of monoids
  or groups in the Crossfilter documentation}.

However, while the presence of the inverse does bring some advantages,
it is also the case that fewer summary statistics have an inverse. For
instance, as was mentioned before, while maximum operator is a valid
monoid, it lacks an inverse and as such does not meet the definition of
a group. Once we summarize some data with the maximum, we cannot
retrieve the second or third highest value - that information is gone
forever. And there are many other operations (such as the convex hull in
Figure \ref{fig:convexhull}) which lack an inverse. As such, when
designing and interactive data visualization library, we need to
carefully weigh the ability to support different kinds of summary
statistics (general, monoids, groups) against the ability to support
wide range of interactions (e.g.~single-group selection vs.~multi-group
selection).

\section{Discussion}\label{discussion}

We hope we have demonstrated that, when designing interactive graphics,
algebraic thinking is invaluable. To support certain interactive
features such as linked selection, the statistics underlying our plots
must possess appropriate algebraic properties. If we want to display the
result of splitting our data into parts and then combining these parts
back together, then we need either monoids or groups. Monoids are for
comparing nested subsets of our data (single-group linked selection)
whereas groups are for comparing disjoint parts of our data (multi-group
linked selection). The properties of these structures ensure that our
figures will be statistically sound, computationally efficient, and
adhere to good design principles.

We do not advocate for exclusive use of groups and monoids. Rather, we
want to highlight the inherent trade-off present in the choice of a
summary statistic. Few summary statistics meet the definition of a
group, but, if we choose such a statistic, then we can do a lot of
things with it. With a monoid, our options are more limited, and with a
non-monoidal statistic, even more so. Like many past authors \citep[such
as][]{tufte2001, tukey1977, wilkinson2012}, we want to emphasize the
importance of thinking critically about the mathematical structure
underlying our graphics. Geometric objects, statistics, and interaction
are all deeply connected, and groups and monoids merely allow us to
carve out a small space of fairly well-behaved visualizations.

On that note, while we have focused on unitality, associativity, and
inverses, it may be useful to consider other algebraic properties as
well. For instance, many of the visual attributes that we use to
represent our data, such as width, height, or area, can only \emph{grow}
in one direction (there is no such thing as ``negative area''). If this
is the case, then our summary statistic should be \emph{monotonic}, such
that if \(x_1 \leq y_1\) and and \(x_2 \leq y_2\) then
\(x_1 \otimes x_2 \leq y_1 \otimes y_2\). Similarly, in many types of
plots, the order of the data points does not matter, however, in some,
such as time-series plots, it does. Thus, it may be important to
determine whether or not the statistics underlying our plots are
\emph{commutative}, such that \(x \otimes y = y \otimes x\). Further
exploration of these properties in relation to interactive graphics may
yield useful insights, however, that is outside of the scope of the
present paper.

On a more practical note, interactive graphics require efficient
pipelines for transforming raw data into summary statistics, and,
currently, we still do not possess a formal framework for doing this. A
decade ago, \citet{wickham2009} noted the lack of such data pipelines,
and, despite some attempts at implementation since then \citep[see
e.g.][]{lawrence2009}, the same sentiment was echoed again recently
\citep{vanderplas2020}. The ultimate goal is to minimize recomputation
during user interaction, and this is where the computational shortcuts
that groups and monoids provide may prove invaluable. We envision a
visualization pipeline structured as a hierarchy of reactive partitions,
where only the downstream (fine) partitions need to be recomputed when
the upstream (coarse) partitions change, but not vice versa. We are
currently exploring this idea in an R package called
\texttt{plotscaper}\footnote{\texttt{https://github.com/bartonicek/plotscaper}}
(note that the project is still under development).

Data visualization, in practice, is rarely about mathematical purity.
Real world data is often flawed or incomplete, and it takes cunning and
experience to dredge up valuable insights from it. Nevertheless, we
believe that simple algebraic thinking offers a valuable foundation.
Concepts like groups and monoids provide a solid framework for reasoning
about graphics, interactive or otherwise. By grounding our thinking in
these concepts, we may be able to offload some of the required cognitive
effort and envision new types of robust and expressive interactive
graphics.

\section{Disclosure Statement}\label{disclosure-statement}

The authors report there are no competing interests to declare.

\bibliographystyle{apalike}
\renewcommand\refname{Bibliography}
\bibliography{references.bib}

\end{document}